### 2014-06-01 ijlyttle/first_pass.Rmd
adds Rmd file where the data are loaded

# First pass through the data

This will be a stream-of-concoiusness walk through the data, a first pass to see what I can see.

## Libraries

At some point, we should make sure we agree on which versions of which packages we propose to use. As a starting point, I can show what I am using:

```{r load_libraries, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggvis)
```

Here's my versions:

```{r session_info}
sessionInfo()
```

I suspect that for publication, `ggplot2` will be more capable than `ggvis`.

# Load the data

```{r load_data}
sets <- c("item", "parent", "school", "scoredItem", "student")

# function to build the file names
fn_build <- function(file_name){

  template <- c("2012.rda", "2012dict.rda")

  file_name %>% 
    vapply(str_join, template, template) %>% 
    file.path("..", "data", .)
}

# load the data
sets %>% fn_build %>% lapply(load, .GlobalEnv)

# clean
rm(fn_build, sets)
```

The dictionaries are stored as named vectors - I would prefer data frames so that I can use the RStudio GUI to examine them.

```{r modify_dictionaries}
# function to convert to data-frames
fn_make_df <- function(named_vector){
  data.frame(
    variable = attr(named_vector, "names"),
    description = named_vector,
    row.names = NULL
  )
}

# there's a clever way to do this, but beyond me for naw
dict_item2012 <- fn_make_df(item2012dict) 
dict_parent2012 <- fn_make_df(parent2012dict) 
dict_school2012 <- fn_make_df(school2012dict) 
dict_scoredItem2012 <- fn_make_df(scoredItem2012dict) 
dict_student2012 <- fn_make_df(student2012dict)

# clean
rm(fn_make_df)
rm(item2012dict, parent2012dict, school2012dict, scoredItem2012dict, student2012dict)
```

Looking at the number of variables in each of the data frames, I wish I knew how to use ggobi...