### 2014-06-10 dicook/Look.Rmd
start on looking at data

PISA
========================================================

Working from Ian's code to read in the data - it is beautiful!

```{r load_libraries, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggvis)
library(maps)
library(ggmap)
```

```{r load_data}
setwd("..")
sets <- c("item", "parent", "school", "scoredItem", "student")

# function to build the file names
fn_build <- function(file_name){

  template <- c("2012.rda", "2012dict.rda")

  file_name %>% 
    vapply(str_join, template, template) %>% 
    file.path(".", "data", .)
}

# load the data
sets %>% fn_build %>% lapply(load, .GlobalEnv)

# clean
#rm(fn_build, sets)
```

```{r modify_dictionaries}
# function to convert to data-frames
fn_make_df <- function(named_vector){
  data.frame(
    variable = attr(named_vector, "names"),
    description = named_vector,
    row.names = NULL
  )
}

# there's a clever way to do this, but beyond me for naw
dict_item2012 <- fn_make_df(item2012dict) 
dict_parent2012 <- fn_make_df(parent2012dict) 
dict_school2012 <- fn_make_df(school2012dict) 
dict_scoredItem2012 <- fn_make_df(scoredItem2012dict) 
dict_student2012 <- fn_make_df(student2012dict)

# clean
#rm(fn_make_df)
#rm(item2012dict, parent2012dict, school2012dict, scoredItem2012dict, student2012dict)
```

Do some basic summaries, to see what's in the data.

```{r whatsthere}
# Schools
dim(dict_school2012)
table(school2012$CNT)
dim(dict_student2012)
length(table(student2012$STIDSTD))
subset(student2012, STIDSTD=="09631")
dict_student2012[,2]
table(student2012$ST04Q01)
length(scoredItem2012dict)
length(item2012dict)
length(parent2012dict)
```

Dizzyingly many variables! (GGobi won't help) Need to map out a plan of attack, questions of interest, and variables to create.

- It will be good to include map information, so we can make maps

- Variables need to be re-organized, many columns represent one level of a categorical variable

- What questions would we have about the data?

- Which variable has the score?

Now create a few variables, to provide examples, and make a few pictures.

```{r usa}
school.usa <- subset(school2012, CNT=="United States of America")
dim(school.usa)
school.usa$GenderPropBoys <- school.usa[,15]/(school.usa[,15]+school.usa[,16])
summary(school.usa$GenderPropBoys)
qplot(GenderPropBoys, data=school.usa)
table(school.usa$SC03Q01)
qplot(GenderPropBoys, SC03Q01, data=school.usa)
summarise(group_by(school.usa, SC03Q01), median(GenderPropBoys, na.rm=T))
qplot(GenderPropBoys, PROPMATH, data=school.usa, ylim=c(0,0.3), xlim=c(0.25, 0.75)) # Ignore extremese
table(school.usa$SC01Q01)
summary(school.usa[,290])
```

Merge data with maps, so it can be plotted on a map. Interesting set of countries in the study - or more what countries are not included! This indicates some inequities.

```{r maps}
student2012.sub <- student2012[,c(1,12)]
student2012.gender <- summarise(group_by(student2012.sub, CNT), PropBoys=length(ST04Q01[ST04Q01=="Male"])/length(ST04Q01))
student2012.gender$CNT <- factor(student2012.gender$CNT, levels=levels(student2012.gender$CNT)[order(student2012.gender$PropBoys)])
qplot(CNT, PropBoys, data=student2012.gender) + coord_flip()
map.CNT <- map()$names
world <- map_data("world")
student2012.gender$CNT <- as.character(student2012.gender$CNT)
student2012.gender$CNT[student2012.gender$CNT=="United States of America"] <- "USA"
student2012.gender.map <- merge(student2012.gender, world, by.x="CNT", by.y="region")
qplot(long, lat, data=student2012.gender.map, group=group, order=order, geom="polygon", fill=PropBoys) + coord_map()
```