### 2014-06-11 alex/initial.Rmd
average hours studied and average scores per country

OECD
========================================================
```{r load, cache=TRUE}
library(ggplot2)
library(dplyr)
setwd("..")
sets <- c("item", "parent", "school", "scoredItem", "student")

# function to build the file names
fn_build <- function(file_name) {

    template <- c("2012.rda", "2012dict.rda")

    file_name %>% vapply(paste0, template, template) %>% file.path(".", "data", 
        .)
}

# load the data
sets %>% fn_build %>% lapply(load, .GlobalEnv)

# function to convert to data-frames
fn_make_df <- function(named_vector) {
    data.frame(variable = attr(named_vector, "names"), description = named_vector, 
        row.names = NULL)
}

# there's a clever way to do this, but beyond me for naw
dict_item2012 <- fn_make_df(item2012dict)
dict_parent2012 <- fn_make_df(parent2012dict)
dict_school2012 <- fn_make_df(school2012dict)
dict_scoredItem2012 <- fn_make_df(scoredItem2012dict)
dict_student2012 <- fn_make_df(student2012dict)
```

Average test scores: each student has 5 scores for each area.  
```{r avg, cache=TRUE}
avg_test = function(subj) {
  apply(student2012[, paste0("PV", 1:5, subj)], 1, mean)
}
subj = c("MATH", "MACC", "MACQ", "MACS", "MACU", "MAPE", "MAPF", "MAPI", "READ", "SCIE")

#rank avg math scores by country
student2012$math = avg_test("MATH")

math_scores = student2012 %>% 
    select(CNT, math) %>% 
    group_by(CNT) %>% 
    summarise(m_math = mean(math)) %>% 
    arrange(m_math)

#rank avg reading scores by country
student2012$read = avg_test("READ")

read_scores = student2012 %>% 
    select(CNT, read) %>% 
    group_by(CNT) %>% 
    summarise(m_read = mean(read)) %>% 
    arrange(m_read)

#rank avg science scores by country
student2012$sci = avg_test("SCIE")

sci_scores = student2012 %>% 
    select(CNT, sci) %>% 
    group_by(CNT) %>% 
    summarise(m_sci = mean(sci)) %>% 
    arrange(m_sci)

print(math_scores)
print(read_scores)
print(sci_scores)
```

I've heard that Chinese students on average spend the most time studying and Finnish students spend the least amount of time.  We can see that Finnish students study very little outside of school.
```{r hours, cache=TRUE}
#removing the stupid attributes from atomic vectors
student2012$MMINS = as.numeric(student2012$MMINS)
student2012$LMINS = as.numeric(student2012$LMINS)
student2012$SMINS = as.numeric(student2012$SMINS)
student2012$ST57Q01 = as.numeric(student2012$ST57Q01)
student2012$ST57Q02 = as.numeric(student2012$ST57Q02)
student2012$ST57Q03 = as.numeric(student2012$ST57Q03)
student2012$ST57Q04 = as.numeric(student2012$ST57Q04)
student2012$ST57Q05 = as.numeric(student2012$ST57Q05)

#all time related vars
dict_student2012[grep("time", dict_student2012[,2], ignore.case=TRUE), ]

OUTHOURS = student2012 %>%
    select(CNT, LMINS, MMINS, OUTHOURS, SMINS, ST57Q01, ST57Q02, ST57Q03, ST57Q04, ST57Q05) %>%
    group_by(CNT) %>%
    summarise(m_OUTHOURS = mean(OUTHOURS, na.rm=T)) %>%
    arrange(m_OUTHOURS)
print(OUTHOURS)
```

We can calculate score per hour to normalize test scores against how much students actually study.
```{r norm, cache=TRUE}
OUTHOURS = OUTHOURS %>% 
  group_by(CNT) %>% 
  left_join(math_scores) %>% 
  left_join(read_scores) %>% 
  left_join(sci_scores) %>%
  mutate(math_hour = m_math / m_OUTHOURS) %>%
  mutate(read_hour = m_read / m_OUTHOURS) %>%
  mutate(sci_hour = m_sci / m_OUTHOURS)

print(OUTHOURS[, c(1,6:8)])
```