### 2014-06-19 alex/2020.Rmd
some stuff

OUTHOURS vs Scores and 20-20 ratio
========================================================

```{r load, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(dplyr)
setwd("..")
sets <- c("item", "parent", "school", "scoredItem", "student")

# function to build the file names
fn_build <- function(file_name) {

    template <- c("2012.rda", "2012dict.rda")

    file_name %>% vapply(paste0, template, template) %>% file.path(".", "data", 
        .)
}

# load the data
sets %>% fn_build %>% lapply(load, .GlobalEnv)

# function to convert to data-frames
fn_make_df <- function(named_vector) {
    data.frame(variable = attr(named_vector, "names"), description = named_vector, 
        row.names = NULL)
}

# there's a clever way to do this, but beyond me for naw
dict_item2012 <- fn_make_df(item2012dict)
dict_parent2012 <- fn_make_df(parent2012dict)
dict_school2012 <- fn_make_df(school2012dict)
dict_scoredItem2012 <- fn_make_df(scoredItem2012dict)
dict_student2012 <- fn_make_df(student2012dict)

rm(fn_make_df, fn_build, sets, item2012dict, parent2012dict, school2012dict, scoredItem2012dict, student2012dict)
```

```{r avg, cache=TRUE, echo=FALSE}
avg_test = function(subj) {
  apply(student2012[, paste0("PV", 1:5, subj)], 1, mean)
}

student2012$math = avg_test("MATH")
student2012$read = avg_test("READ")
student2012$sci = avg_test("SCIE")
```

While looking at my favorite country with a city named Tirana, I noticed that something strange was happening.  No other country appears to have this.
```{r albania, echo=FALSE, cache=TRUE}
library(ggplot2)
#peak at 54 hours
qplot(OUTHOURS, data = subset(student2012, CNT == "Albania"))
```

Not much correlation between scores and out of school study hours
```{r outhours, cache=TRUE, tidy=FALSE}
qplot(x = OUTHOURS, y = math, 
      data = student2012 %>% 
        select(OUTHOURS, math) %>% 
        filter(!is.na(OUTHOURS)), 
      alpha = I(0.1)) + stat_smooth()
qplot(x = OUTHOURS, y = read, 
      data = student2012 %>% 
        select(OUTHOURS, read) %>% 
        filter(!is.na(OUTHOURS)), 
      alpha = I(0.1) ) + stat_smooth()
qplot(x = OUTHOURS, y = sci, 
      data = student2012 %>% 
        select(OUTHOURS, sci) %>% 
        filter(!is.na(OUTHOURS)), 
      alpha = I(0.1) ) + stat_smooth()
```

```{r 2020, cache=TRUE, echo=FALSE, eval=FALSE}
##UNUSED
#cutoffs for top 20/bottom 20 for each country
best_worst = student2012 %>% 
  select(CNT, math, read, sci, OUTHOURS) %>% 
  group_by(CNT) %>% 
  summarise(math_low = quantile(math, probs = seq(0, 1, 0.2))[2],
            math_high = quantile(math, probs = seq(0, 1, 0.2))[5],
            read_low = quantile(read, probs = seq(0, 1, 0.2))[2],
            read_high = quantile(read, probs = seq(0, 1, 0.2))[5],
            sci_low = quantile(sci, probs = seq(0, 1, 0.2))[2],
            sci_high = quantile(sci, probs = seq(0, 1, 0.2))[5])
head(best_worst)
```

Mean scores in the top 20 and bottom 20 percent for each nation
```{r 2020_1, cache=TRUE, tidy=FALSE}
library(reshape2)
best_worst = student2012 %>% 
  select(CNT, math, read, sci) %>% 
  group_by(CNT) %>% 
  summarise(math_high = mean(math[math > quantile(math, probs = seq(0, 1, 0.2))[5]]),
            math_low = mean(math[math < quantile(math, probs = seq(0, 1, 0.2))[2]]),
            sci_high = mean(sci[sci > quantile(sci, probs = seq(0, 1, 0.2))[5]]),
            sci_low = mean(sci[sci < quantile(sci, probs = seq(0, 1, 0.2))[2]]),
            read_high = mean(read[read > quantile(read, probs = seq(0, 1, 0.2))[5]]),
            read_low = mean(read[read < quantile(read, probs = seq(0, 1, 0.2))[2]])) %>%
  mutate(math_gap = math_high - math_low,
         sci_gap = sci_high - sci_low,
         read_gap = read_high - read_low) %>%
  mutate(avg_gap = (math_gap + sci_gap + read_gap)/3 ) %>%
  arrange(desc(avg_gap))
head(best_worst)

best_worst$CNT = factor(best_worst$CNT, 
                        levels = best_worst$CNT[order(best_worst$avg_gap)])
qplot(x = avg_gap, y = CNT, data = best_worst)

qplot(x = value, y = CNT, data = best_worst %>% 
        select(CNT, math_gap, read_gap, sci_gap) %>% 
        melt(id.vars = "CNT"), color = variable)
```

Compare differences in the top 20 and bottom 20 math scores for each nation.
```{r math, cache=TRUE}
student2012$ST69Q02 = as.numeric(student2012$ST69Q02)
student2012$ST70Q02 = as.numeric(student2012$ST70Q02)

math2020 = student2012 %>%
            select(CNT, math, OUTHOURS, ST69Q02, ST70Q02, #minutes studying, class periods
                   ST44Q01, ST44Q03, ST44Q04,ST44Q05, ST44Q07, ST44Q08, #attributions to failure
                   ST15Q01, ST19Q01) %>% #job status for parents
            group_by(CNT) %>%
            mutate(top20 = math > quantile(math, probs = seq(0, 1, 0.2))[5]) %>%
            filter(math < quantile(math, probs = seq(0, 1, 0.2))[2] | 
                   math > quantile(math, probs = seq(0, 1, 0.2))[5])
```

```{r math2, eval=TRUE}
library(ggplot2)

#Attributions to Failure - Not Good at Maths Problems
qplot(x = ST44Q01, fill = top20, data = subset(math2020, !is.na(ST44Q01)), geom = "bar", position = "fill")

#Attributions to Failure - Teacher Did Not Explain Well
qplot(x = ST44Q03, fill = top20, data = subset(math2020, !is.na(ST44Q03)), geom = "bar", position = "fill")

#Attributions to Failure - Bad Guesses
qplot(x = ST44Q04, fill = top20, data = subset(math2020, !is.na(ST44Q04)), geom = "bar", position = "fill")

#Attributions to Failure - Material Too Hard
qplot(x = ST44Q05, fill = top20, data = subset(math2020, !is.na(ST44Q05)), geom = "bar", position = "fill")

#Attributions to Failure - Teacher Didnt Get Students Interested
qplot(x = ST44Q07, fill = top20, data = subset(math2020, !is.na(ST44Q07)), geom = "bar", position = "fill")

#Attributions to Failure - Unlucky
qplot(x = ST44Q08, fill = top20, data = subset(math2020, !is.na(ST44Q08)), geom = "bar", position = "fill")

#Father Current Job Status
qplot(x = ST19Q01, fill = top20, data = subset(math2020, !is.na(ST19Q01)), position = "fill")

#mother current job status
qplot(x = ST15Q01, fill = top20, data = subset(math2020, !is.na(ST15Q01)), position = "fill")

#hours studied
qplot(x = OUTHOURS, fill = top20, data = subset(math2020, !is.na(OUTHOURS)), position = "fill")
```