### 2014-06-21 ijlyttle/factor_explorer.Rmd
gets student explorer going

---
title: "factor_explorer"
runtime: shiny
output: html_document
---

## Intro

ggvis is amlmost there, but not quite, so I'll use ggplot.

## Libraries

At some point, we should make sure we agree on which versions of which packages we propose to use. As a starting point, I can show what I am using:

```{r load_libraries, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(shiny)
library(magrittr)

# 2014-06-20
# for what it's worth, I'm running the latest versions of RStudio (0.98.932 )and 
# github versions of "rstudio/shiny", "rstudio/rmarkdown", "yihui/knitr"
```

Here's my versions:

```{r session_info}
sessionInfo()
```

```{r load_data}
sets <- c("item", "parent", "school", "scoredItem", "student")

# function to build the file names
fn_build <- function(file_name){

  template <- c("2012.rda", "2012dict.rda")

  file_name %>% 
    vapply(str_join, template, template) %>% 
    file.path("..", "data", .)
}

# load the data
sets %>% fn_build %>% lapply(load, .GlobalEnv)

# clean
rm(fn_build, sets)
```

The dictionaries are stored as named vectors - I would prefer data frames so that I can use the RStudio GUI to examine them.

```{r modify_dictionaries}
# function to convert to data-frames
fn_make_df <- function(named_vector){
  data.frame(
    variable = attr(named_vector, "names"),
    description = named_vector,
    row.names = NULL
  )
}

# there's a clever way to do this, but beyond me for naw
dict_item2012 <- fn_make_df(item2012dict) 
dict_parent2012 <- fn_make_df(parent2012dict) 
dict_school2012 <- fn_make_df(school2012dict) 
dict_scoredItem2012 <- fn_make_df(scoredItem2012dict) 
dict_student2012 <- fn_make_df(student2012dict)

# clean
rm(fn_make_df)
#rm(item2012dict, parent2012dict, school2012dict, scoredItem2012dict, student2012dict)
```

## Shiny app

We are told that we can put a shiny app into an Rmarkdown document - let's try that.

The thought is that we could use this (I wish we had another couple weeks) to discover the interesting relationships, then turn those into posters.

```{r shiny_variable_lists}
var_names_student <- dict_student2012$variable
names(var_names_student) <- 
  dict_student2012$description %>%
  str_replace_all("\x92", "'") # gets rid of funny characters"

var_names_student_subject <- 
  var_names_student[c(
    seq(501, 505), # math
    seq(541, 545), # reading
    seq(546, 551)  # science
  )]

var_names_student_type <- 
  lapply(student2012, class)

var_names_student_factor <- 
  var_names_student_type[var_names_student_type == "factor"] %>%
  names

var_names_factor <- var_names_student[var_names_student %in% var_names_student_factor] 
```

```{r shiny_inputs, echo = FALSE}
selectInput(
  "subject",
  label = "Subject", 
  choices = var_names_student_subject
)

selectInput(
  "country",
  label = "Countries",
  choices = levels(student2012$CNT),
  selected = c("United States of America", "Canada"),
  multiple = TRUE
)

selectInput(
  "x_factor",
  label = "X-axis factor",
  choices = var_names_factor,
  selected = "ST04Q01"
)

selectInput(
  "fill_factor",
  label = "Fill factor",
  choices = var_names_factor,
  selected = "ST26Q03"
)
```

```{r shiny_transform, echo = FALSE}
rct_data <- reactive({
  df <- 
    student2012 %>%
    filter(CNT %in% input$country) 
})

val_get_name <- function(x, lst){
  names(lst[lst == x])[[1]]
}

rct_count <- reactive({
  df <-
    rct_data() %>%
    group_by()
})

rct_labels <- reactive({
  list(
    x = input$x_factor %>% val_get_name(var_names_factor),
    y = input$subject %>% val_get_name(var_names_student_subject),
    fill = input$fill_factor %>% val_get_name(var_names_factor)
  )
})

renderPlot({
  ggplot(
    data = rct_data()
  ) + 
  geom_violin(
    aes_string(
      x = input$x_factor, 
      y = input$subject),
      scale = "width"
  ) + 
  geom_boxplot(
    aes_string(
      x = input$x_factor,
      y = input$subject,
      fill = input$fill_factor
    )  
  ) + 
  facet_grid(CNT ~ .) + 
  scale_x_discrete(name = rct_labels()[["x"]]) +  
  scale_y_continuous(limits = c(0, 1000), name = rct_labels()[["y"]]) +
  scale_fill_discrete(name = rct_labels()[["fill"]]) 
})
```