### 2014-06-22 dicook/possessions.Rmd
looking at possessions, age

Exploring possessions, and other demographics
========================================================

```{r load_libraries, message=FALSE}
library(stringr)
library(reshape2)
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggvis)
library(maps)
library(ggmap)
library(rworldmap)
library (grid)    
library (scales)	
```

```{r load_data, echo=FALSE}
setwd("..")
sets <- c("item", "parent", "school", "scoredItem", "student")

# function to build the file names
fn_build <- function(file_name){

  template <- c("2012.rda", "2012dict.rda")

  file_name %>% 
    vapply(str_join, template, template) %>% 
    file.path(".", "data", .)
}

# load the data
sets %>% fn_build %>% lapply(load, .GlobalEnv)

# clean
#rm(fn_build, sets)
```

```{r modify_dictionaries, echo=FALSE}
# function to convert to data-frames
fn_make_df <- function(named_vector){
  data.frame(
    variable = attr(named_vector, "names"),
    description = named_vector,
    row.names = NULL
  )
}

# there's a clever way to do this, but beyond me for naw
dict_item2012 <- fn_make_df(item2012dict) 
dict_parent2012 <- fn_make_df(parent2012dict) 
dict_school2012 <- fn_make_df(school2012dict) 
dict_scoredItem2012 <- fn_make_df(scoredItem2012dict) 
dict_student2012 <- fn_make_df(student2012dict)

# clean
#rm(fn_make_df)
#rm(item2012dict, parent2012dict, school2012dict, scoredItem2012dict, student2012dict)
```

```{r usefulfn, echo=FALSE}
extractPolygons <- function(shapes) {

  dframe <- ldply(1:length(shapes@polygons), function(i) {
    ob <- shapes@polygons[[i]]@Polygons
    dframe <- ldply(1:length(ob), function(j) {
      x <- ob[[j]]
      co <- x@coords
      data.frame(co, order=1:nrow(co), group=j)
    })
    dframe$region <- i
    dframe$name <- shapes@polygons[[i]]@ID
    dframe
  })
  # construct a group variable from both group and polygon:
  dframe$group <- interaction(dframe$region, dframe$group)

  dframe
}

# To get a blank background on map
new_theme_empty <- theme_bw()
new_theme_empty$line <- element_blank()
new_theme_empty$rect <- element_blank()
new_theme_empty$strip.text <- element_blank()
new_theme_empty$axis.text <- element_blank()
new_theme_empty$plot.title <- element_blank()
new_theme_empty$axis.title <- element_blank()
new_theme_empty$plot.margin <- structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit")
```

```{r countpossessions, warning=FALSE, message=FALSE}
student2012.sub <- student2012[, c(1:7, 44:57, 61:66, seq(501, 550, 5))]
colnames(student2012.sub)[1] <- "name"
for (i in 8:21) {
  student2012.sub[,i] <- as.character(student2012.sub[,i])
  student2012.sub[is.na(student2012.sub[,i]), i] <- ""
  student2012.sub[student2012.sub[,i]=="Yes",i] <- "1"
  student2012.sub[student2012.sub[,i]=="No",i] <- "0"
  student2012.sub[,i] <- as.numeric(student2012.sub[,i])
}
for (i in 22:26) {
  student2012.sub[,i] <- as.character(student2012.sub[,i])
  student2012.sub[is.na(student2012.sub[,i]), i] <- ""
  student2012.sub[student2012.sub[,i]=="None",i] <- "0"
  student2012.sub[student2012.sub[,i]=="One",i] <- "1"
  student2012.sub[student2012.sub[,i]=="Two",i] <- "2"
  student2012.sub[student2012.sub[,i]=="Three or more",i] <- "3"
  student2012.sub[,i] <- as.numeric(student2012.sub[,i])
}
student2012.sub[,27] <- as.character(student2012.sub[,27])
student2012.sub[is.na(student2012.sub[,27]), 27] <- ""
student2012.sub[student2012.sub[,27]=="0-10 books ",27] <- "0"
student2012.sub[student2012.sub[,27]=="11-25 books ",27] <- "1"
student2012.sub[student2012.sub[,27]=="26-100 books ",27] <- "2"
student2012.sub[student2012.sub[,27]=="101-200 books ",27] <- "3"
student2012.sub[student2012.sub[,27]=="201-500 books ",27] <- "4"
student2012.sub[student2012.sub[,27]=="More than 500 books",27] <- "5"
student2012.sub[,27] <- as.numeric(student2012.sub[,27])
student2012.sub$numposs <- apply(student2012.sub[,8:21], 1, sum, na.rm=T)
student2012.sub$numedposs <- apply(student2012.sub[,22:27], 1, sum, na.rm=T)
```

```{r plotstuff, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
qplot(numposs, PV1MATH, data=student2012.sub, alpha=I(0.1), geom=c("point", "smooth"))
qplot(numedposs, PV1MATH, data=student2012.sub, alpha=I(0.1), geom=c("point", "smooth"))
student2012.sub$PV1MATH <- as.numeric(student2012.sub$PV1MATH)
student2012.sub$PV1MACC <- as.numeric(student2012.sub$PV1MACC)
student2012.sub$PV1MACQ <- as.numeric(student2012.sub$PV1MACQ)
student2012.sub$PV1MACS <- as.numeric(student2012.sub$PV1MACS)
student2012.sub$PV1MACU <- as.numeric(student2012.sub$PV1MACU)
student2012.sub$PV1MAPE <- as.numeric(student2012.sub$PV1MAPE)
student2012.sub$PV1MAPF <- as.numeric(student2012.sub$PV1MAPF)
student2012.sub$PV1MAPI <- as.numeric(studen=t2012.sub$PV1MAPI)
student2012.sub$PV1READ <- as.numeric(student2012.sub$PV1READ)
student2012.sub$PV1SCIE <- as.numeric(student2012.sub$PV1SCIE)
student2012.sub.summary <- summarise(group_by(student2012.sub[,c(1,8:39)], name), math=mean(PV1MATH, na.rm=T), 
          read=mean(PV1READ, na.rm=T), science=mean(PV1SCIE, na.rm=T),
          poss = mean(numposs, na.rm=T), edposs = mean(numedposs, na.rm=T))
qplot(poss, math, data=student2012.sub.summary) + geom_smooth(method="lm", se=F)
qplot(poss, read, data=student2012.sub.summary) + geom_smooth(method="lm", se=F)
qplot(poss, science, data=student2012.sub.summary) + geom_smooth(method="lm", se=F)
head(student2012.sub.summary[order(student2012.sub.summary$math, decreasing=T),], 20) 
# Vietnam is the outlier, few poss, high scores
lm(math~poss, data=student2012.sub.summary)
# Each increase in one possession tends to increase math score by 25 points (poss=7, 12)
qplot(edposs, math, data=student2012.sub.summary) + geom_smooth(method="lm", se=F)
qplot(edposs, read, data=student2012.sub.summary) + geom_smooth(method="lm", se=F)
qplot(edposs, science, data=student2012.sub.summary) + geom_smooth(method="lm", se=F)
lm(math~edposs, data=student2012.sub.summary)
# Each increase in one possession tends to increase math score by 13 points (edposs=5, 14)
qplot(edposs, poss, data=student2012.sub.summary)
```

```{r plotstuff2, fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
student2012.sub.summary <- summarise(group_by(student2012.sub[,c(1,8,28,38)], name), 
              math05=mean(PV1MATH[numposs<6], na.rm=T), math06=mean(PV1MATH[numposs==6], na.rm=T), 
              math07=mean(PV1MATH[numposs==7], na.rm=T), math08=mean(PV1MATH[numposs==8], na.rm=T),
              math09=mean(PV1MATH[numposs==9], na.rm=T), math10=mean(PV1MATH[numposs==10], na.rm=T),
              math11=mean(PV1MATH[numposs==11], na.rm=T), math12=mean(PV1MATH[numposs==12], na.rm=T),
              math13=mean(PV1MATH[numposs==13], na.rm=T), math14=mean(PV1MATH[numposs==14], na.rm=T))
student2012.sub.summary.m <- melt(student2012.sub.summary)
student2012.sub.summary.m$variable <- substr(student2012.sub.summary.m$variable, 5, 6) 
student2012.sub.summary.m$variable <- as.numeric(student2012.sub.summary.m$variable)
qplot(variable, value, data=student2012.sub.summary.m, group=name, geom=c("point","smooth"), se=F) + facet_wrap(~name, ncol=8)
```

```{r age, fig.width=8, fig.height=12, warning=FALSE, message=FALSE}
summary(student2012$ST06Q01)
student2012$ST06Q01 <- as.numeric(student2012$ST06Q01)
student2012.age.iqr <- summarise(group_by(student2012[,c(1,14)], CNT), q0=min(ST06Q01, na.rm=T),
                                 q25=quantile(ST06Q01, 0.25, na.rm=T), q50=median(ST06Q01, na.rm=T), 
                                 q75=quantile(ST06Q01, 0.75, na.rm=T), q100=max(ST06Q01, na.rm=T), 
                                 count=length(ST06Q01))
student2012.age.iqr$CNT <- factor(student2012.age.iqr$CNT, 
      levels=student2012.age.iqr$CNT[order(student2012.age.iqr$q50)])
ggplot(data=student2012.age.iqr) + ylab("Age") + xlab("") + 
  geom_point(aes(x=CNT, y=q50, size=count)) + geom_segment(aes(x=CNT, xend=CNT, y=q0, yend=q25)) +
  geom_segment(aes(x=CNT, xend=CNT, y=q75, yend=q100)) +
  coord_flip() + theme(legend.position="none")
```

```{r agemath, fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
student2012.sub <- student2012[, c(1, 14, 501)]
colnames(student2012.sub)[1] <- "name"
student2012.sub$PV1MATH <- as.numeric(student2012.sub$PV1MATH)
student2012.sub$ST06Q01 <- as.numeric(student2012.sub$ST06Q01)
student2012.sub.summary <- summarise(group_by(student2012.sub, name), math=mean(PV1MATH, na.rm=T), 
          age=mean(ST06Q01, na.rm=T))
qplot(age, math, data=student2012.sub.summary)
# Age at start school does not seem to make a difference on country level
# qplot(ST06Q01, PV1MATH, data=student2012, alpha=I(0.1), xlab="Age", ylab="Math")
student2012.sub.summary <- summarise(group_by(student2012.sub, name), math4=mean(PV1MATH[ST06Q01==4], na.rm=T), 
          math5=mean(PV1MATH[ST06Q01==5], na.rm=T), math6=mean(PV1MATH[ST06Q01==6], na.rm=T), 
          math7=mean(PV1MATH[ST06Q01==7], na.rm=T), math8=mean(PV1MATH[ST06Q01>7], na.rm=T))
student2012.sub.summary.m <- melt(student2012.sub.summary)
student2012.sub.summary.m$variable <- substr(student2012.sub.summary.m$variable, 5, 5)
student2012.sub.summary.m$variable <- as.numeric(student2012.sub.summary.m$variable)
qplot(variable, value, data=student2012.sub.summary.m, group=name, geom=c("point","smooth"), se=F) + facet_wrap(~name, ncol=8)
# Looking by country is important: generally starting school early leads to better scores
```