### 2014-06-23 alex/maps.Rmd
simple map

Maps
========================================================
```{r load, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(plyr)
library(dplyr)
setwd("..")
sets <- c("item", "parent", "school", "scoredItem", "student")

# function to build the file names
fn_build <- function(file_name) {

    template <- c("2012.rda", "2012dict.rda")

    file_name %>% vapply(paste0, template, template) %>% file.path(".", "data", 
        .)
}

# load the data
sets %>% fn_build %>% lapply(load, .GlobalEnv)

# function to convert to data-frames
fn_make_df <- function(named_vector) {
    data.frame(variable = attr(named_vector, "names"), description = named_vector, 
        row.names = NULL)
}

dict_item2012 <- fn_make_df(item2012dict)
dict_parent2012 <- fn_make_df(parent2012dict)
dict_school2012 <- fn_make_df(school2012dict)
dict_scoredItem2012 <- fn_make_df(scoredItem2012dict)
dict_student2012 <- fn_make_df(student2012dict)

rm(fn_make_df, fn_build, sets, item2012dict, parent2012dict, school2012dict, scoredItem2012dict, student2012dict)
```

```{r time, warning=FALSE, message=FALSE}
#Relevant variables for hours studied
hours = student2012 %>%
    select(CNT, LMINS, MMINS, OUTHOURS, SMINS, ST57Q01, ST57Q02, ST57Q03, ST57Q04, ST57Q05,
           PV1MATH, PV1SCIE, PV1READ)

hours$MMINS = as.numeric(hours$MMINS)
hours$LMINS = as.numeric(hours$LMINS)
hours$SMINS = as.numeric(hours$SMINS)
hours$ST57Q01 = as.numeric(hours$ST57Q01)
hours$ST57Q02 = as.numeric(hours$ST57Q02)
hours$ST57Q03 = as.numeric(hours$ST57Q03)
hours$ST57Q04 = as.numeric(hours$ST57Q04)
hours$ST57Q05 = as.numeric(hours$ST57Q05)
hours$PV1MATH = as.numeric(hours$PV1MATH)
hours$PV1READ = as.numeric(hours$PV1READ)
hours$PV1SCIE = as.numeric(hours$PV1SCIE)

#change country names
hours$CNT = as.character(hours$CNT)
hours$CNT[hours$CNT=="Serbia"] <- "Republic of Serbia"
hours$CNT[hours$CNT=="Korea"] <- "South Korea"
hours$CNT[hours$CNT=="Chinese Taipei"] <- "Taiwan"
hours$CNT[hours$CNT=="Slovak Republic"] <- "Slovakia"
hours$CNT[hours$CNT=="Russian Federation"] <- "Russia"
hours$CNT[hours$CNT=="Perm(Russian Federation)"] <- "Russia"
hours$CNT[hours$CNT=="Hong Kong-China"] <- "Hong Kong S.A.R."
hours$CNT[hours$CNT=="China-Shanghai"] <- "China"
hours$CNT[hours$CNT=="China-Macau"] <- "China"
hours$CNT[hours$CNT=="Connecticut (USA)"] <- "United States of America"
hours$CNT[hours$CNT=="Florida (USA)"] <- "United States of America"
hours$CNT[hours$CNT=="Massachusetts (USA)"] <- "United States of America"

#calculate means by CNT
hours = hours %>%
    group_by(CNT) %>%
    summarise(m_OUTHOURS = mean(OUTHOURS, na.rm=T), math = mean(PV1MATH, na.rm=T),
              sci = mean(PV1SCIE, na.rm=T), read = mean(PV1READ, na.rm=T)) %>%
    arrange(m_OUTHOURS) %>% mutate(mscore = (math + sci + read)/3)

hours$CNT = factor(hours$CNT,
                      levels = hours$CNT[order(hours$m_OUTHOURS)])
names(hours)[1] = "name"
print(hours)
p1 = qplot(name, m_OUTHOURS, data = hours, color = m_OUTHOURS, size = mscore) + coord_flip()
```

```{r usefulfn, echo=FALSE, warning=FALSE, message=FALSE}
extractPolygons <- function(shapes) {

  dframe <- ldply(1:length(shapes@polygons), function(i) {
    ob <- shapes@polygons[[i]]@Polygons
    dframe <- ldply(1:length(ob), function(j) {
      x <- ob[[j]]
      co <- x@coords
      data.frame(co, order=1:nrow(co), group=j)
    })
    dframe$region <- i
    dframe$name <- shapes@polygons[[i]]@ID
    dframe$area <- shapes@polygons[[i]]@area
    dframe
  })
  # construct a group variable from both group and polygon:
  dframe$group <- interaction(dframe$region, dframe$group)

  dframe
}

# To get a blank background on map
new_theme_empty <- theme_bw()
new_theme_empty$line <- element_blank()
new_theme_empty$rect <- element_blank()
new_theme_empty$strip.text <- element_blank()
new_theme_empty$axis.text <- element_blank()
new_theme_empty$plot.title <- element_blank()
new_theme_empty$axis.title <- element_blank()
new_theme_empty$plot.margin <- structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit")
```

```{r map, warning=FALSE, message=FALSE}
library(maps)
library(rworldmap)
world <- getMap(resolution = "low")
world.polys <- extractPolygons(world)
hours.map = left_join(hours, world.polys)

p2 = ggplot(data=world.polys) + geom_path(aes(x=X1, y=X2, order=order, group=group), colour=I("grey70")) + 
  geom_polygon(data=hours.map, aes(x=X1, y=X2, order=order, group=group, fill=m_OUTHOURS)) +
  new_theme_empty + theme(legend.position="none")

p1 = ggplotGrob(p1)
p2 + annotation_custom(grob = p1, xmin = -40, xmax = 80, ymin = -110, ymax = 10)
```