### 2014-06-23 ijlyttle/factor_explorer.Rmd
adds color bands

---
title: "factor_explorer"
runtime: shiny
output: html_document
---

## Intro

ggvis is amlmost there, but not quite, so I'll use ggplot.

## Libraries

At some point, we should make sure we agree on which versions of which packages we propose to use. As a starting point, I can show what I am using:

```{r load_libraries, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(shiny)
library(magrittr)

# 2014-06-20
# for what it's worth, I'm running the latest versions of RStudio (0.98.932 )and 
# github versions of "rstudio/shiny", "rstudio/rmarkdown", "yihui/knitr", "hadley/tidyr"
```

Here's my versions:

```{r session_info}
sessionInfo()
```

```{r load_data}
#sets <- c("item", "parent", "school", "scoredItem", "student")
sets <- "student"

# function to build the file names
fn_build <- function(file_name){

  template <- c("2012.rda", "2012dict.rda")

  file_name %>% 
    vapply(str_join, template, template) %>% 
    file.path("..", "data", .)
}

# load the data
sets %>% fn_build %>% lapply(load, .GlobalEnv)

# clean
rm(fn_build, sets)
```

The dictionaries are stored as named vectors - I would prefer data frames so that I can use the RStudio GUI to examine them.

```{r modify_dictionaries}
# function to convert to data-frames
fn_make_df <- function(named_vector){
  data.frame(
    variable = attr(named_vector, "names"),
    description = named_vector,
    row.names = NULL
  )
}

# there's a clever way to do this, but beyond me for naw
# dict_item2012 <- fn_make_df(item2012dict) 
# dict_parent2012 <- fn_make_df(parent2012dict) 
# dict_school2012 <- fn_make_df(school2012dict) 
# dict_scoredItem2012 <- fn_make_df(scoredItem2012dict) 
dict_student2012 <- fn_make_df(student2012dict)

# clean
rm(fn_make_df)
#rm(item2012dict, parent2012dict, school2012dict, scoredItem2012dict, student2012dict)
```

## Shiny app

We are told that we can put a shiny app into an Rmarkdown document - let's try that.

The thought is that we could use this (I wish we had another couple weeks) to discover the interesting relationships, then turn those into posters.

```{r make_factors}
fn_make_factor <- function(vec){  
  vec <- factor(vec, levels = sort(unique(vec)))
}

# hacky - fix when time
student2012$ST01Q01 <- fn_make_factor(student2012$ST01Q01)
student2012$ST02Q01 <- fn_make_factor(student2012$ST02Q01)
student2012$ST06Q01 <- fn_make_factor(student2012$ST06Q01)
student2012$ST21Q01 <- fn_make_factor(student2012$ST21Q01)
student2012$ST115Q01 <- fn_make_factor(student2012$ST115Q01)

```

```{r shiny_variable_lists}
var_names_student <- dict_student2012$variable
names(var_names_student) <- 
  dict_student2012$description %>%
  str_replace_all("\x92", "'") # gets rid of funny characters"

var_names_student_subject <- 
  var_names_student[c(
    seq(501, 505), # math
    seq(541, 545), # reading
    seq(546, 550)  # science
  )]

var_names_student_type <- 
  lapply(student2012, class)

var_names_student_factor <- 
  var_names_student_type[var_names_student_type == "factor"] %>%
  names

var_names_factor <- var_names_student[var_names_student %in% var_names_student_factor] 
```

```{r shiny_inputs, echo = FALSE}
selectInput(
  "subject",
  label = "Subject", 
  choices = var_names_student_subject
)

selectInput(
  "country",
  label = "Countries",
  choices = levels(student2012$CNT),
  selected = c("United States of America", "Peru", "China-Shanghai"),
  multiple = TRUE
)

selectInput(
  "x_factor",
  label = "X-axis factor",
  choices = var_names_factor,
  selected = "ST27Q03"
)

selectInput(
  "fill_factor",
  label = "Fill factor",
  choices = var_names_factor,
  selected = "ST28Q01"
)
```

```{r shiny_transform, echo = FALSE}
rct_data <- reactive({
  df <- 
    student2012 %>%
    filter(CNT %in% input$country) 
})

val_get_name <- function(x, lst){
  names(lst[lst == x])[[1]]
}

rct_count <- reactive({

  # hacky hacky hacky  
  str_symbols <- c("CNT", input$x_factor, input$fill_factor) 
  symbols <- str_symbols %>% lapply(as.symbol)

  df_data <- rct_data()

  if (nrow(df_data) == 0)
    return(NULL)

#   # get counts by category
   df <- 
     df_data[, str_symbols] %>%
     regroup(symbols) %>%
     summarize(n = n())

  # get counts by country  
  df_cnt <-
    rct_data() %>%
    group_by(CNT) %>%
    summarize(n_country = n(), y_placeholder = 0)

  # join, get proportion by country for each category
  df_new <- 
    left_join(df, df_cnt, by = "CNT") %>%
    mutate(prop = n / n_country)

  df_new
})

observe({  
  print(rct_count())
  print("")
#  print(str(rct_count()))
#  print("")
#  print(summary(rct_count()))
})

rct_labels <- reactive({
  list(
    x = input$x_factor %>% val_get_name(var_names_factor),
    y = input$subject %>% val_get_name(var_names_student_subject),
    fill = input$fill_factor %>% val_get_name(var_names_factor)
  )
})

renderPlot({
  ggplot(
    data = rct_data()
  ) + 
  geom_rect(
    aes(fill = CNT),
    xmin = -Inf, xmax = Inf, 
    ymin = -Inf, ymax = Inf,
    alpha = 0.5,
    data = data.frame(CNT = unique(rct_data()$CNT))
  ) + 
  geom_violin(
    aes_string(
      x = input$x_factor, 
      y = input$subject),
      scale = "width"
  ) + 
#   geom_boxplot(
#     aes_string(
#       x = input$x_factor,
#       y = input$subject,
#       group = input$fill_factor
#     ),
#     fill = "white",
#     alpha = 1,
#   ) + 
  geom_boxplot(
    aes_string(
      x = input$x_factor,
      y = input$subject,
      alpha = input$fill_factor
    ),
    fill = "blue"
  ) + 
  facet_grid(CNT ~ .) + 
  scale_x_discrete(name = rct_labels()[["x"]]) +  
  scale_y_continuous(limits = c(0, 1000), name = rct_labels()[["y"]]) +
  scale_alpha_discrete(range = c(0.1, 0.9), na.value = 0, name = rct_labels()[["fill"]]) +  
  scale_fill_brewer(type = "seq", palette = "BuGn", guide = FALSE) 
})

renderPlot({
  ggplot(
    data = rct_data()
  ) +
  geom_rect(
    aes(fill = CNT),
    xmin = -Inf, xmax = Inf, 
    ymin = -Inf, ymax = Inf,
    alpha = 0.5,
    data = data.frame(CNT = unique(rct_data()$CNT))
  ) + 
  geom_bar(
    aes_string(
      x = input$x_factor,
      group = input$fill_factor
    ),
    position = "dodge", 
    fill = "white", 
    alpha = 1,
    color = "black"
  ) +   
  geom_bar(
    aes_string(
      x = input$x_factor,
      alpha = input$fill_factor
    ),
    position = "dodge", 
    fill = "blue"
  ) +   
  facet_grid(CNT ~ ., scales = "free_y") + 
  scale_x_discrete(name = rct_labels()[["x"]]) +  
  scale_y_continuous(name = "number of students") +
  scale_alpha_discrete(range = c(0.1, 0.9), na.value = 0, name = rct_labels()[["fill"]]) +  
  scale_fill_brewer(type = "seq", palette = "BuGn", guide = FALSE) 
})
```