### 2014-06-27 alex/albania.Rmd
albania

Equality for all: Only in student2012.sub
=============
```{r load, echo=FALSE, results='hide', warning = FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
setwd("..")
sets <- c("school", "student")

# function to build the file names
fn_build <- function(file_name) {
    template <- c("2012.rda", "2012dict.rda")
    file_name %>% vapply(paste0, template, template) %>% file.path(".", "data", 
        .)
}

# load the data
sets %>% fn_build %>% lapply(load, .GlobalEnv)

# function to convert to data-frames
fn_make_df <- function(named_vector) {
    data.frame(variable = attr(named_vector, "names"), description = named_vector, 
        row.names = NULL)
}

dict_school2012 <- fn_make_df(school2012dict)
dict_student2012 <- fn_make_df(student2012dict)
rm(fn_make_df, fn_build, sets, school2012dict, student2012dict)

student2012$math = as.numeric(student2012[, "PV1MATH"])
student2012$read = as.numeric(student2012[, "PV1READ"])
student2012$sci = as.numeric(student2012[, "PV1SCIE"])
```

```{r vars}
student2012.sub = student2012 %>% select(CNT, SCHOOLID, math, read, sci, OUTHOURS, STIDSTD, ST04Q01, ST11Q01, ST11Q02, ST15Q01, ST19Q01, ST26Q01, ST26Q02, ST26Q03, ST26Q04, ST26Q05, ST26Q06, ST26Q07, ST26Q08, ST26Q09, ST26Q10, ST26Q11, ST26Q12, ST26Q13, ST26Q14, ST27Q01, ST27Q02, ST27Q03, ST27Q04, ST27Q05, ST28Q01, OECD)

for (i in 13:26) {
    student2012.sub[, i] <- as.character(student2012.sub[, i])
    student2012.sub[is.na(student2012.sub[, i]), i] <- ""
    student2012.sub[student2012.sub[, i] == "Yes", i] <- "1"
    student2012.sub[student2012.sub[, i] == "No", i] <- "0"
    student2012.sub[, i] <- as.numeric(student2012.sub[, i])
}
for (i in 27:31) {
    student2012.sub[, i] <- as.character(student2012.sub[, i])
    student2012.sub[is.na(student2012.sub[, i]), i] <- ""
    student2012.sub[student2012.sub[, i] == "None", i] <- "0"
    student2012.sub[student2012.sub[, i] == "One", i] <- "1"
    student2012.sub[student2012.sub[, i] == "Two", i] <- "2"
    student2012.sub[student2012.sub[, i] == "Three or more", i] <- "3"
    student2012.sub[, i] <- as.numeric(student2012.sub[, i])
}
student2012.sub[, 32] <- as.character(student2012.sub[, 32])
student2012.sub[is.na(student2012.sub[,32]), 32] <- ""
student2012.sub[student2012.sub[, 32] == "0-10 books ", 32] <- "0"
student2012.sub[student2012.sub[, 32] == "11-25 books ", 32] <- "1"
student2012.sub[student2012.sub[, 32] == "26-100 books ", 32] <- "2"
student2012.sub[student2012.sub[, 32] == "101-200 books ", 32] <- "3"
student2012.sub[student2012.sub[, 32] == "201-500 books ", 32] <- "4"
student2012.sub[student2012.sub[, 32] == "More than 500 books", 32] <- "5"
student2012.sub[, 32] <- as.numeric(student2012.sub[, 32])

#aggregate possessions
student2012.sub$numposs <- apply(student2012.sub[, 13:26], 1, sum, na.rm = T)
student2012.sub$numedposs <- apply(student2012.sub[, 27:31], 1, sum, na.rm = T)

#separate albania
student2012.sub$OECD = as.character(student2012.sub$OECD)
student2012.sub[student2012.sub$CNT == "Albania", ]$OECD = "Albania"
```

```{r possessions, echo=FALSE}
qplot()
```